Parsing Postfix log files

### 2007-12-06 doc/logparser.tex
Now saving the size of mails delivered.
Replaced regexs with regexes.

XXX NOW SAVING THE SIZE OF MAILS\@.

@@

@@

Perl compiles the original \regex{} into an internal representation,
optimising the \regex{} to improve the speed of matching, but this
compilation and optimisation takes CPU time; far more CPU time, in fact,
than the actual matching takes.  Perl automatically caches static
\regexs{}, but dynamic \regexs{} need to be explicitly compiled and cached.
Graph~\refwithpage{normal regex vs discard regex} shows execution times
with and without caching the \regex{}.  Caching the compiled \regexs{} is
obviously far more efficient; graph~\refwithpage{normal regex vs discarded
regex factor} shows the percentage execution time increase when not caching
each \regex{}.

Caching the compiled \regexs{} is quite simple, and is the single most
effective optimisation implemented in the parser.

@@

    \item [Pre-process the file]  The new algorithm leverages the knowledge
        gained while writing rules and performs a large number of
        substitutions on the input lines, replacing commonly occurring
        variable terms (e.g.\ email addresses, \IP{} addresses, the
        standard start of rejection messages, etc.) with \regex{} keywords
        which the parser will expand when it loads the rule (see the
        \regex{} entry in section~\refwithpage{rule attributes}).  The
        purpose of this phase is to utilise the existing knowledge to
        create more accurate \regexs{}.  The new log lines are written to a
        temporary file, which all subsequent stages use instead of the
        original input file.

@@

    \item [Build regexs]  Contiguous sequences of \texttt{.*} in the newly
        reassembled \regexs{} are collapsed to a single \texttt{.*}; any
        resulting duplicate \regexs{} are combined, and their counters
        summed.  If the frequency of a \regex{} is lower than the threshold
        supplied by the user the \regex{} is discarded.\footnote{This is a
        second, independent threshold, but once again 1\% of input lines is
        a good starting point; obviously the threshold depends on the
        number and type of input lines.}  The new \regexs{} are printed for
        the user to add to the database, either as new rules or merged into
        the \regexs{} of existing rules; the counter for each \regex{} is
        also printed, giving the user an indication of how many of the
        input lines that \regex{} should match.  Discarding \regexs{} will
        result in some of the input lines not being matched; this utility
        should be run again once the unmatched lines have been isolated by
        running the parser, including the new \regexs{}, with the same
        input which produced the original set of unparsed lines.

@@

A second utility is also provided which reads a list of new \regexs{} and
the input given to the first utility.  It tries to match each input line
against each \regex{}, counting the number of lines which match each
\regex{}, warning the user if an input line is matched by more than one
\regex{}, and additionally warning if an input line is not matched by any
\regex{}.  It displays a summary of how many input lines each \regex{}
matched, comparing it to the expected number of matches; this provides the
user with an easy method of checking if the \regexs{} produced by the first
utility are correctly matching the input lines they are based upon.  A
future version of this utility will also group input lines by \regex{}, so
the user can tweak the \regexs{} if required.

@@

    \item The topic of rule efficiency was discussed next, covering the
        effects of caching compiled \regexs{} and optimal ordering of
        rules, with reference to the graphs in
        appendix~\refwithpage{graphs}.

@@

Warnings are issued for any lines which are not parsed; no warnings are
issued for unparsed lines while testing with the 93 test log files, so it
can be safely concluded that there are zero false negatives.  False
positives are harder to quantify: short of examining each of the 60,721,709
log lines and determining which \regex{} parsed it, there is no way to be
sure that every line was parsed by the correct \regex{}, making it
impossible to quantify the false positive rate.  The author is confident
that the false positive rate is extremely low, approaching if not reaching
zero, because the \regexs{} used are as specific as possible and care has
been taken to ensure that there are no unintentionally overlapping
\regexs{}.

@@

    \item [Log analyser --- throughput monitor] This utility tracks the
        number of events which occurred over a particular time and warns if
        the frequency of events passes a certain threshold.  It's designed
        to provide real time alerts when dictionary attacks, mail loops or
        similar problems occur. It doesn't attempt to extract or save data,
        correlate log lines, or any of the more advanced tasks described in
        this document, because it is not designed to do so.  The user must
        construct the \regexs{} to match significant input lines before it
        will parse any logs.  This program was not reviewed in further
        detail because its aims are so far removed from the aims of this
        project.

@@

Graphs are an excellent means of displaying data, transforming a
meaningless stream of numbers into an easily comprehensible form, where
anomalies and patterns are immediately obvious.  These graphs are used to
illustrate the topics discussed in section~\refwithpage{rule efficiency}.
The graphs in the first section cover parser scalability, demonstrating
that performance scales linearly with input size.  The impact of rule
ordering is shown in section~\refwithpage{rule ordering graphs}, and the
anomalous dips and peaks apparent in some graphs are explained.  The third
group of graphs vividly shows the huge impact that caching compiled
\regexs{} has on parser performance.  This section concludes with a
breakdown of the rule hits accumulated during a single test run of the
parser.

@@

\subsection{Caching regexs}

The following graphs show the impact that not caching compiled \regexs{}
has on parser performance: on typical log files the execution time when not
caching compiled \regexs{} is 500--600\% of the execution time when
caching; reversing the perspective shows that cached execution time is
merely 17--20\% of non-cached execution time.  Caching compiled \regexs{}
is probably the single most effective optimisation possible in the parser's
implementation.

@@

The distribution of log lines across rules when there is a mail loop is
much different and the average number of rules consulted per log line is
much lower; this results in far fewer \regex{} compilations per line than
when there isn't a mail loop, and a correspondingly decreased execution
time.  The increases in execution time when not caching \regexs{} for
different combinations of log files are summarised in the table below:

@@

The graphs presented in this section illustrate the topics discussed in
section~\refwithpage{rule efficiency}.  The first collection of graphs are
about parser scalability, showing the linear relationship between execution
time and input size.  Section~\refwithpage{rule ordering graphs} relates to
the effect of rule ordering on execution time, and the unexpected
consequences of specific inputs.  The necessity of caching compiled
\regexs{} is attested to by the third group of graphs, where the difference
between caching and discarding is staggering.  This section concluded with
a breakdown of the rule hits accumulated during a single test run of the
parser.

>>>>

@@

    \item [size] The size of the mail; it will only be present for
        delivered mails.

@@

Perl compiles the original \regex{} into an internal representation,
optimising the \regex{} to improve the speed of matching, but this
compilation and optimisation takes CPU time; far more CPU time, in fact,
than the actual matching takes.  Perl automatically caches static
\regexes{}, but dynamic \regexes{} need to be explicitly compiled and
cached.  Graph~\refwithpage{normal regex vs discard regex} shows execution
times with and without caching the \regex{}.  Caching the compiled
\regexes{} is obviously far more efficient; graph~\refwithpage{normal regex
vs discarded regex factor} shows the percentage execution time increase
when not caching each \regex{}.

Caching the compiled \regexes{} is quite simple, and is the single most
effective optimisation implemented in the parser.

@@

    \item [Pre-process the file]  The new algorithm leverages the knowledge
        gained while writing rules and performs a large number of
        substitutions on the input lines, replacing commonly occurring
        variable terms (e.g.\ email addresses, \IP{} addresses, the
        standard start of rejection messages, etc.) with \regex{} keywords
        which the parser will expand when it loads the rule (see the
        \regex{} entry in section~\refwithpage{rule attributes}).  The
        purpose of this phase is to utilise the existing knowledge to
        create more accurate \regexes{}.  The new log lines are written to
        a temporary file, which all subsequent stages use instead of the
        original input file.

@@

    \item [Build regexes]  Contiguous sequences of \texttt{.*} in the newly
        reassembled \regexes{} are collapsed to a single \texttt{.*}; any
        resulting duplicate \regexes{} are combined, and their counters
        summed.  If the frequency of a \regex{} is lower than the threshold
        supplied by the user the \regex{} is discarded.\footnote{This is a
        second, independent threshold, but once again 1\% of input lines is
        a good starting point; obviously the threshold depends on the
        number and type of input lines.}  The new \regexes{} are printed
        for the user to add to the database, either as new rules or merged
        into the \regexes{} of existing rules; the counter for each
        \regex{} is also printed, giving the user an indication of how many
        of the input lines that \regex{} should match.  Discarding
        \regexes{} will result in some of the input lines not being
        matched; this utility should be run again once the unmatched lines
        have been isolated by running the parser, including the new
        \regexes{}, with the same input which produced the original set of
        unparsed lines.

@@

A second utility is also provided which reads a list of new \regexes{} and
the input given to the first utility.  It tries to match each input line
against each \regex{}, counting the number of lines which match each
\regex{}, warning the user if an input line is matched by more than one
\regex{}, and additionally warning if an input line is not matched by any
\regex{}.  It displays a summary of how many input lines each \regex{}
matched, comparing it to the expected number of matches; this provides the
user with an easy method of checking if the \regexes{} produced by the
first utility are correctly matching the input lines they are based upon.
A future version of this utility will also group input lines by \regex{},
so the user can tweak the \regexes{} if required.

@@

    \item The topic of rule efficiency was discussed next, covering the
        effects of caching compiled \regexes{} and optimal ordering of
        rules, with reference to the graphs in
        appendix~\refwithpage{graphs}.

@@

Warnings are issued for any lines which are not parsed; no warnings are
issued for unparsed lines while testing with the 93 test log files, so it
can be safely concluded that there are zero false negatives.  False
positives are harder to quantify: short of examining each of the 60,721,709
log lines and determining which \regex{} parsed it, there is no way to be
sure that every line was parsed by the correct \regex{}, making it
impossible to quantify the false positive rate.  The author is confident
that the false positive rate is extremely low, approaching if not reaching
zero, because the \regexes{} used are as specific as possible and care has
been taken to ensure that there are no unintentionally overlapping
\regexes{}.

@@

    \item [Log analyser --- throughput monitor] This utility tracks the
        number of events which occurred over a particular time and warns if
        the frequency of events passes a certain threshold.  It's designed
        to provide real time alerts when dictionary attacks, mail loops or
        similar problems occur. It doesn't attempt to extract or save data,
        correlate log lines, or any of the more advanced tasks described in
        this document, because it is not designed to do so.  The user must
        construct the \regexes{} to match significant input lines before it
        will parse any logs.  This program was not reviewed in further
        detail because its aims are so far removed from the aims of this
        project.

@@

Graphs are an excellent means of displaying data, transforming a
meaningless stream of numbers into an easily comprehensible form, where
anomalies and patterns are immediately obvious.  These graphs are used to
illustrate the topics discussed in section~\refwithpage{rule efficiency}.
The graphs in the first section cover parser scalability, demonstrating
that performance scales linearly with input size.  The impact of rule
ordering is shown in section~\refwithpage{rule ordering graphs}, and the
anomalous dips and peaks apparent in some graphs are explained.  The third
group of graphs vividly shows the huge impact that caching compiled
\regexes{} has on parser performance.  This section concludes with a
breakdown of the rule hits accumulated during a single test run of the
parser.

@@

\subsection{Caching regexes}

The following graphs show the impact that not caching compiled \regexes{}
has on parser performance: on typical log files the execution time when not
caching compiled \regexes{} is 500--600\% of the execution time when
caching; reversing the perspective shows that cached execution time is
merely 17--20\% of non-cached execution time.  Caching compiled \regexes{}
is probably the single most effective optimisation possible in the parser's
implementation.

@@

The distribution of log lines across rules when there is a mail loop is
much different and the average number of rules consulted per log line is
much lower; this results in far fewer \regex{} compilations per line than
when there isn't a mail loop, and a correspondingly decreased execution
time.  The increases in execution time when not caching \regexes{} for
different combinations of log files are summarised in the table below:

@@

The graphs presented in this section illustrate the topics discussed in
section~\refwithpage{rule efficiency}.  The first collection of graphs are
about parser scalability, showing the linear relationship between execution
time and input size.  Section~\refwithpage{rule ordering graphs} relates to
the effect of rule ordering on execution time, and the unexpected
consequences of specific inputs.  The necessity of caching compiled
\regexes{} is attested to by the third group of graphs, where the
difference between caching and discarding is staggering.  This section
concluded with a breakdown of the rule hits accumulated during a single
test run of the parser.
